MAESTROeX: A Massively Parallel Low Mach Number Astrophysical Solver

### 2019-05-23 MAESTROeX_I/intro.tex
add cite to JOSS AMReX
some comments

\section{Introduction} \label{sec:intro}
Many astrophysical flows are highly subsonic; sound waves carry sufficiently little energy that they do not significantly affect the convective dynamics of the system.
In many of these flows, modeling long-time convective dynamics are of interest, and numerical approaches based on compressible hydrodynamics are intractable, even on modern supercomputers.
One approach to this problem is to use low Mach number models.
In a low Mach number approach, sound waves are eliminated from the governing equations while retaining compressibilitiy effects due to, e.g., nuclear energy release, compositional changes, and thermal diffusion.
When the Mach number (the ratio of the characteristic fluid velocity over the characteristic sound speed; Ma $= U/c$) is small,
the resulting system can be numerically integrated with much larger time steps than a compressible model.
Specifically, the time step increase is at least a factor of $\sim 1/{\rm Ma}$ larger.
Each time step is more computationally expensive due to the presence of additional linear solves,
but for many problems of interest the overall gains in efficiency can easily be an order of magnitude or more.

@@

In this paper, we present new algorithmic methodology that improves upon Paper V in a number of ways.
First, the overall temporal algorithm has been greatly simplified without compromising accuracy.
The key design decisions were to eliminate the splitting of the velocity into average and perturbational components, 
and also to replace the hydrodynamic evolution of the base state with a predictor-corrector approach.
Not only does this greatly simplify the dynamics of the base
state, but this treatment is more amenable to higher-order multiphysics coupling strategies
based on method-of-lines integration.  
In particular, schemes based on spectral deferred corrections \citep{dutt2000spectral} have been used to generate high-order temporal integrators for problems of reactive flow \citep{dutt2000spectral} and low Mach number combustion \citep{pazner2016high,nonaka2018conservative}.
Second, we explore the effects of alternative spatial mapping routines for coupling the base state and the Cartesian grid state in spherical problems.
Finally, the code has been completely ported to a new software framework suitable for modern manycore supercomputers.
We examine the performance of our new MAESTROeX implementation in the new C++/F90 AMReX public software library \cite{AMReX}.
MAESTROeX uses MPI+OpenMP parallelism and scales well to over 10,000 MPI processes, with each MPI process supporting tens of threads.
The resulting code is publicly available on GitHub ({\tt https://github.com/AMReX-Astro/MAESTROeX}),
uses the Starkiller-Astro microphysics libraries ({\tt https://github.com/starkiller-astro/Microphysics}),
and the AMReX software library for block-structured adaptive mesh refinement ({\tt https://github.com/AMReX-Codes/amrex}).

>>>>

\section{Introduction} \label{sec:intro}
Many astrophysical flows are highly subsonic.  In this regime,
sound waves carry sufficiently little energy that they do not
significantly affect the convective dynamics of the system.  In many
of these flows, modeling long-time convective dynamics are of
interest, and numerical approaches based on compressible hydrodynamics
are intractable, even on modern supercomputers.  One approach to this
problem is to use low Mach number models.  In a low Mach number
approach, sound waves are eliminated from the governing equations
while retaining compressibilitiy effects due to, e.g., nuclear energy
release, stratification, compositional changes, and thermal diffusion.  When the Mach
number (the ratio of the characteristic fluid velocity over the
characteristic sound speed; Ma $= U/c$) is small, the resulting system
can be numerically integrated with much larger time steps than a
compressible model.  Specifically, the time step increase is at least
a factor of $\sim 1/{\rm Ma}$ larger.  Each time step is more
computationally expensive due to the presence of additional linear
solves, but for many problems of interest the overall gains in
efficiency can easily be an order of magnitude or more.

@@

In this paper, we present new algorithmic methodology that improves upon Paper V in a number of ways.
First, the overall temporal algorithm has been greatly simplified without compromising accuracy.
The key design decisions were to eliminate the splitting of the velocity into average and perturbational components, 
and also to replace the hydrodynamic evolution of the base state with a predictor-corrector approach.
Not only does this greatly simplify the dynamics of the base
state, but this treatment is more amenable to higher-order multiphysics coupling strategies
based on method-of-lines integration.  
In particular, schemes based on spectral deferred corrections have been used to generate high-order temporal integrators for problems of reactive flow \citep{dutt2000spectral} and low Mach number combustion \citep{pazner2016high,nonaka2018conservative}.
Second, we explore the effects of alternative spatial mapping routines for coupling the base state and the Cartesian grid state in spherical problems.
Finally, the code has been completely ported to a new software framework suitable for modern manycore and GPU-based supercomputers.
We examine the performance of our new MAESTROeX implementation in the new C++/F90 AMReX public software library \cite{AMReX,AMReX_JOSS}.
MAESTROeX uses MPI+OpenMP parallelism and scales well to over 10,000 MPI processes, with each MPI process supporting tens of threads.
The resulting code is publicly available on GitHub ({\tt https://github.com/AMReX-Astro/MAESTROeX}),
uses the Starkiller-Astro microphysics libraries ({\tt https://github.com/starkiller-astro/Microphysics}),
and the AMReX software library for block-structured adaptive mesh refinement ({\tt https://github.com/AMReX-Codes/amrex}).
